\section{Additional finite examples}

\subsection{Accumulating transport defect}
The verified family used in Figure~2 begins with two source macrostates. Each independently exposed binary target distinction forces a separate split inside the inherited class, producing
\[
|Q_0|=2,\qquad |Q_1^{\min}|=2^m+1,
\]
and therefore $\Delta_{\#}=2^m-1$. This construction is a sharp finite witness showing that repair burden can grow rapidly when target operations independently expose previously suppressed coordinates. It is not presented as an empirical ecological scaling law.

\subsection{Coherent replacement histories}
When two declared replacement routes induce the same complete carried terminal map, they share one inherited terminal partition. The source-relative repair and transport defect are then route independent, so no historical mode is retained.

\subsection{Incoherent replacement histories}
When two routes induce different complete carried terminal maps, at least one terminal configuration receives incompatible inherited meanings. A route-free carried law is impossible. Grouping histories by equality of their complete maps gives the minimum immutable context, after which exact source-relative refinement is performed within the augmented state space.

\subsection{Interpretation of the plant--pollinator witness}
The main-text example uses the smallest verified local split: inherited labels $(0,0,1)$ are repaired to $(0,1,2)$. The additional state records substitute-pollinator response capacity. The example is intentionally finite and diagnostic: it demonstrates how a target-only management action can reverse a restoration priority without claiming to estimate the ecological transition structure from field data.
